---
title: "Project: Spatial Bros"
description: "Exploring Network-Constrained Spatial Point Analysis"
author: "Spatial Bros"
date: "27/03/2023"
number-sections: true
categories: ["network-constrained spatial point analysis"]
title-block-banner: true
---

# Network Constrained Spatial Point Pattern Analysis

## Project Overview

Network constrained Spatial Point Patterns Analysis (NetSPAA) is a collection of spatial point patterns analysis methods special developed for analysing spatial point event occurs on or alongside network.

The spatial point event can be locations of childcare centre for example.

The network, on the other hand can be a road network for example.

More information with regards to data sources used for this project can be found at the [proposal](https://spatialbros.netlify.app/proposal.html#data-sources)

## Installing the R packages

In this project, x packages will be used

```{r}
pacman::p_load(sp, sf, rgdal, spNetwork, tmap)
```

## Data Import and Prepration

### Melbourne Road Network

```{r}
road_network <- st_read("data/geospatial/Roads Network", layer = "road-corridors")
```

### Melbourne Pedestrian Network

```{r}
pedestrian_network <- st_read("data/geospatial/Pedestrian Network/pedestrian-network.geojson")
```

### Tram Network

```{r}
tram_network <- st_read("data/geospatial/Trams Network", layer = "PTV_METRO_TRAM_ROUTE")
```

### Localities

```{r}
localities = st_read("data/geospatial/Localities", layer = "UCL_2021_AUST_GDA2020") 
melbourne_localities <- localities[localities$UCL_NAME21 == "Melbourne", ]

```

### Local Government Areas

```{r}
local_government_areas = st_read("data/geospatial/Local Government Areas", layer = "LGA_2022_AUST_GDA2020")
melbourne_local_government_areas = local_government_areas[local_government_areas$LGA_NAME22 == "Melbourne",]
```
